A Memory-Bounded, Deterministic and Terminating Semantics for the Synchronous Programming Language~\CEU

### 2018-05-28 06-appendix.tex
(*) final version

\lemxdetout*
\begin{proof}
  The lemma is vacuously true if~$\delta$ cannot be advanced by~$\out$
  transitions.  Suppose that is not the case and let~$\delta=\<p,n,e>$,
  $\delta_1=\<p_1,n_1,e_1>$ and~$\delta_2=\<p_2,n_2,e_2>$.  Then, there are
  two possibilities.
  \begin{case}
  \item $e\ne\nil$.  Both transitions are applications of~\R{push}.
    Hence~$p_1=p_2=\bcast(p,e)$, $n_1=n_2=n+1$, and~$e_1=e_2=\nil$.
    %%
  \item $e=\nil$.  Both transitions are applications of~\R{pop}.
    Hence $p_1=p_2=p$, $n_1=n_2=n-1$, and~$e_1=e_2=\nil$.\qedhere
  \end{case}
\end{proof}

@@

    Hence~$p_1=p_2=\ceu{\Nop}$ and~$e_1=e_2=\nil$.

@@

    Hence~$p_1=p_2=\ceu{\CanRun(n)}$ and~$e_1=e_2=e'$.

@@

      Hence~$p_1=p_2=p'$ and~$e_1=e_2=\nil$.

@@

      Hence~$p_1=p_2=p''$ and~$e_1=e_2=\nil$.

@@

      Hence~$p_1=p_2=\ceu{\Break}$ and~$e_1=e_2=\nil$.

@@

      Hence~$p_1=\ceu{p_1';p''}=\ceu{p_2';p''}=p_2$

@@

    Hence~$p_1=p_2=\ceu{p'\AtLoop{p'}}$ and~$e_1=e_2=\nil$.

@@

      Hence~$p_1=p_2=\ceu{\Loop{p''}}$ and~$e_1=e_2=\nil$.

@@

      Hence~$p_1=\ceu{p_1'\AtLoop{p''}}=\ceu{p_2'\AtLoop{p''}}=p_2$

@@

    Hence~$p_1=p_2=\ceu{{p'}\AtAnd{(\CanRun(n);\,p'')}}$

@@

        Hence~$p_1=p_2=\ceu{\clear(p'');\Break}$ and~$e_1=e_2=\nil$.

@@

        Hence~$p_1=\ceu{{p_1'}\And{p''}}=\ceu{{p_2'}\And{p''}}=p_2$

@@

        Hence~$p_1=p_2=\ceu{\clear(p');\Break}$ and~$e_1=e_2=\nil$.

@@

        Hence~$p_1=\ceu{{p'}\And{p_1''}}=\ceu{{p'}\And{p_2''}}=p_2$

@@

    Hence~$p_1=p_2=\ceu{{p'}\AtOr{(\CanRun(n);\,p'')}}$

@@

        Hence~$p_1=p_2=\clear(p'')$ and~$e_1=e_2=\nil$.

@@

        Hence~$p_1=p_2=\clear(p')$ and~$e_1=e_2=\nil$.

@@

  \item $p=\ceu{p';\,p''}$.
    %%
    There are three subcases.
    \begin{case}
    \item\label{thm.x.term-nst-*.seq-nop} $p'=\ceu{\Nop}$.
      %%
      Then, by~\R{seq-nop} and by the induction hypothesis, there is
      a~$\delta'$ such that
      \[
        \<\ceu{\Nop;\,p''},n,\nil>
        \nst[1]\<p'',n,e>\nst[*]\delta'_\Hnst\,.
      \]
      %%
    \item\label{thm.x.term-nst-*.seq-brk} $p'=\ceu{\Break}$.
      %%
      Then, by~\R{seq-brk},
      \[
        \<\ceu{\Break;\,p''},n,\nil>\nst[1]\<\ceu{\Break},n,\nil>_\Hnst\,.
      \]
      %%
    \item\label{thm.x.term-nst-*.seq-adv}
      $p'\ne\ceu{\Nop},\ceu{\Break}$.
      %%
      Then, by the induction hypothesis, there are~$p_1'$ and~$e$ such that
      \[
        \<p',n,\nil>\nst[*]\<p_1',n,e>_\Hnst\,.
      \]
      By item~\ref{lem.x.props-nst-i.a} of Lemma~\ref{lem.x.props-nst-i},
      \begin{equation}
        \label{thm.x.term-nst-*.seq-adv.eq1}
        \<\ceu{p';\,p''},n,\nil>\nst[*]\<\ceu{p_1';\,p''},n,e>\,.
      \end{equation}
      It remains to be shown that~$\<\ceu{p_1';\,p''},n,e>$ is
      nested-irreducible.  There are four possibilities following from the
      fact that the simpler~$\<p_1',n,e>$ is nested-irreducible.
      %%
      \begin{case}
      \item $e\ne\nil$.  Then, by the definition of~$\Hnst$,
        description~$\<\ceu{p_1';\,p''},n,e>$ is nested-irreducible.
        %%
      \item $p_1'=\ceu{\Nop}$.
        %%
        From~\eqref{thm.x.term-nst-*.seq-adv.eq1},
        \[
          \<\ceu{p';\,p''},n,\nil>\nst[*]\<\ceu{\Nop;\,p''},n,e>\,.
        \]
        From this point on, this case is similar to
        Case~\ref{thm.x.term-nst-*.seq-nop}.
        %%
      \item $p_1'=\ceu{\Break}$.
        %%
        From~\eqref{thm.x.term-nst-*.seq-adv.eq1},
        \[
          \<\ceu{p';\,p''},n,\nil>\nst[*]\<\ceu{\Break;\,p''},n,e>\,.
        \]
        From this point on, this case is similar to
        Case~\ref{thm.x.term-nst-*.seq-brk}.
        %%
      \item$\isblocked(p_1',n)$.
        %%
        Then, by definition,
        \[
          \isblocked(\ceu{p_1';p''},n)=\isblocked(p_1',n)=\mathit{true}\,.
        \]
        Hence, from~\eqref{thm.x.term-nst-*.seq-adv.eq1} and by the
        definition~$\Hnst$, description~$\<\ceu{p_1';\,p''},n,e>$ is
        nested-irreducible.
      \end{case}
    \end{case}

  \item\label{thm.x.term-nst-*.loop}
    $p=\ceu{\Loop{p'}}$.
    %%
    Then, by item~\ref{ass.x.syn-rest.loop} of
    Assumption~\ref{ass.x.syn-rest},
    \begin{equation}\label{thm.x.term-nst-*.loop-expd.eq1}
      \<\ceu{\Loop{p'}},n,\nil>\nst[*]\<p_1',n,e>\,,
    \end{equation}
    for some~$e$ and~$p_1'$ such that either~$p_1'=\ceu{\Break\AtLoop{p'}}$
    or~$\isblocked(p_1',n)$.
    \begin{case}
    \item$p_1'=\ceu{\Break\AtLoop{p'}}$.
      %%
      From~\eqref{thm.x.term-nst-*.loop-expd.eq1}, by~\R{loop-brk},
      \begin{align*}
        \<\ceu{\Loop{p'}},n,\nil>
        &\nst[*]\<\ceu{\Break\AtLoop{p'}},n,e>\\
        &\nst[1]\<\ceu{\Nop},n,e>_\Hnst\,.
      \end{align*}
      %%
    \item$\isblocked(p_1',n)$.  Hence,
      from~\eqref{thm.x.term-nst-*.loop-expd.eq1} and by the definition
      of~$\Hnst$, $\<p_1',n,e>_\Hnst$.
    \end{case}

@@

\lemxrankout*
\begin{proof}\strut
  Let~$\delta=\<p,n,e>$, $\delta'=\<p',n',e'>$,
  $\rank(\delta)=\<i,j>$, and~$\rank(\delta')=\<i',j'>$.
  \begin{enumerate:a}
  \item Suppose~$\<p,n,e>\outpush\<p',n',e'>$.  Then, by~\R{push},
    $e\ne\nil$, $p'=\bcast(p,e)$, $n'=n+1$, and~$e'=\nil$.  By
    Definition~\ref{def.x.rank}, $j=n+1$, as~$e\ne\nil$, and~$j'=n+1$,
    as~$e'=\nil$ and~$n'=n+1$; hence~$j=j'$.
    %%
    It remains to be shown that~$i=i'$:
    \begin{align*}
      i&=\pot(p,e)
         \tag*{by Definition~\ref{def.x.rank}}\\
       &=\pot'(\bcast(p,e))
         \tag*{by Definition~\ref{def.x.pot}}\\
       &=\pot'(p')
         \tag*{since~$p'=\bcast(p,e)$}\\
       &=\pot'(\bcast(p',\nil))
         \tag*{by definition of~$\bcast$}\\
       &=\pot'(\bcast(p',e'))
         \tag*{since~$e'=\nil$}\\
       &=\pot(p',e')
         \tag*{by Definition~\ref{def.x.pot}}\\
       &=i'
         \tag*{by Definition~\ref{def.x.rank}}
    \end{align*}
    Therefore, $\<i,j>=\<i',j'>$.

  \item Suppose~$\<p,n,e>\outpop\<p',n',e'>$.  Then, by~\R{pop}, $p=p'$,
    $n>0$, $n'=n-1$, and~$e=e'=\nil$.
    %%
    By Definition~\ref{def.x.pot}, $\pot(\bcast(p,e))=\pot(\bcast(p',e'))$;
    hence~$i=i'$.  And by Definition~\ref{def.x.rank}, $j=n$, as~$e=\nil$,
    and~$j'=n-1$, as~$e'=\nil$ and~$n'=n-1$; hence~$j>j'$.
    Therefore, ${\<i,j>}>{\<i',j'>}$.\qedhere
  \end{enumerate:a}
\end{proof}

\lemxranknst*
\begin{proof}
  We proceed by induction on the structure of~$\nst$ derivations.
  Let~$\delta=\<p,n,e>$, $\delta'=\<p',n',e'>$, $\rank(\delta)=\<i,j>$,
  and~$\rank(\delta')=\<i',j'>$.  By the hypothesis of the lemma, there is
  a derivation~$\pi$ such that
  \[
    \pi\Vdash\<p,n,e>\nst\<p',n',e'>\,.
  \]
  By definition of~$\nst$, $e=\nil$ and $n=n'$.  Depending on the structure
  of program~$p$, there are~11 possibilities.  In each one of them we show
  that~$\rank(\delta)\ge\rank(\delta')$.
  \begin{case}
  \item$p=\ceu{\Mem(id)}$.
    %%
    Then~$\pi$ is an instance of~\R{mem}.  Hence~$p'=\ceu{\Nop}$
    and~$e'=\nil$.  Thus $\rank(\delta)=\rank(\delta')=\<0,n>$.

  \item$p=\ceu{\EmitInt(e_1)}$.
    %%
    Then~$\pi$ is an instance of~\R{emit-int}.  Hence~$p'=\ceu{\CanRun}$
    and~$e'=e_1\ne\nil$.
    Thus
    \[
      {\rank(\delta)={\<1,n>}}>{\<0,n+1>=\rank(\delta')}\,.
    \]

  \item$p=\ceu{\CanRun(n)}$.
    %%
    Then~$\pi$ is an instance of~\R{can-run}.  Hence~$p'=\ceu{\Nop}$
    and~$e'=\nil$.  Thus
    \[
      \rank(\delta)=\rank(\delta')=\<0,n>\,.
    \]

  \item$p=\ceu{\IfElse{p}{p_1}{p_2}}$.
    %%
    There are two subcases.
    \begin{case}
    \item\label{lem.x.rank-nst.if-true}$\eval(\ceu{\Mem(\Id)})$.
      %%
      Then~$\pi$ is an instance of~\R{if-true}.  Hence~$p'=\ceu{p_1}$
      and~$e'=\nil$.  Thus
      \begin{align*}
        \rank(\delta)&=\<\max\{pot'(p_1),pot'(p_2)\},n>\\
                     &\ge\<\pot'(p_1),n>=\rank(\delta')\,.
      \end{align*}
      %%
    \item$\lnot\eval(\ceu{\Mem(\Id)})$.
      %%
      Similar to Case~\ref{lem.x.rank-nst.if-true}.
    \end{case}

  \item$p=\ceu{p_1;\,p_2}$.
    %%
    There are three subcases.
    \begin{case}
    \item\label{lem.x.rank-nst.seq-nop}$p_1=\ceu{\Nop}$.
      %%
      Then~$\pi$ is an instance of~\R{seq-nop}.
      Hence~$p'=p_2$ and~$e'=\nil$.  Thus
      \begin{align*}
        \rank(\delta)&=\<\pot'(p_1)+\pot'(p_2),n>\\
                     &\ge\<pot'(p_2),n>=\rank(\delta')\,.
      \end{align*}
      %%
    \item\label{lem.x.rank-nst.seq-brk}$p_1=\ceu{\Break}$.
      %%
      Then~$\pi$ is an instance of~\R{seq-brk}.
      Hence~$p'=p_1$ and~$e'=\nil$.  Thus
      \[
        \rank(\delta)=\rank(\delta')=\<0,n>\,.
      \]
      %%
    \item\label{lem.x.rank-nst.seq-adv}$p_1\ne\ceu{\Nop},\ceu{\Break}$.
      %%
      Then~$\pi$ is an instance of~\R{seq-adv}.  Hence there is a
      derivation~$\pi'$ such that
      \[
        \pi'\Vdash\<p_1,n,\nil>\nst\<p_1',n,e_1'>\,,
      \]
      for some~$p_1'$ and~$e_1'$.  Thus~$p'=p_1';p_2$ and~$e'=e_1'$.  By the
      induction hypothesis,
      \begin{equation}
        \label{lem.x.rank-nst.seq-adv.eq1}
        \rank(\<p_1,n,\nil>)\ge\rank(\<p_1',n,e_1'>)\,.
      \end{equation}
      There are two subcases.
      \begin{case}
      \item$e'=\nil$
        %%
        Then
        \begin{align*}
          \rank(\delta)&=\<\pot'(p_1)+\pot'(p_2),n>\enspace\text{and}\\
          \rank(\delta')&=\<\pot'(p_1')+\pot'(p_2),n>\,.
        \end{align*}
        By~\eqref{lem.x.rank-nst.seq-adv.eq1}, $\pot'(p_1)\ge\pot'(p_1')$.
        Thus
        \[
          \rank(\delta)\ge\rank(\delta')\,.
        \]
        %%
      \item$e'\ne\nil$.
        %%
        Then~$\pi'$ contains one application of~\R{emit-int}, which consumes
        one~$\ceu{\EmitInt(e')}$ statement from~$p_1$ and implies
        $\pot'(p_1)>\pot'(p_1')$.  Thus
        \begin{align*}
          \rank(\delta)&=\<\pot'(p_1)+\pot'(p_2),n>\\
                       &>\<\pot'(p_1')+\pot'(p_2),n+1>=\rank(\delta')\,.
        \end{align*}
      \end{case}
    \end{case}

  \item\label{lem.x.rank-nst.loop-expd}$p=\ceu{\Loop{p_1}}$.
    %%
    Then~$\pi$ is an instance of~\R{loop-expd}.
    Hence~$p'=\ceu{p_1\AtLoop{p_1}}$ and~$e'=\nil$.
    %%
    By item~\ref{ass.x.syn-rest.loop} of Assumption~\ref{ass.x.syn-rest},
    all execution paths of~$p_1$ contain at least one occurrence
    of~$\ceu{\Break}$ or~$\ceu{\AwaitExt}$.  Thus, by condition~(\dag) in
    Definition~\ref{def.x.pot},
    \[
      \rank(\delta)=\rank(\delta')=\<\pot'(p_1),n>\,.
    \]

  \item$p=\ceu{{p_1}\AtLoop{p_2}}$.
    %%
    There are three subcases.
    \begin{case}
    \item$p_1=\ceu{\Nop}$.
      %%
      Similar to Case~\ref{lem.x.rank-nst.seq-nop}.
      %%
    \item$p_1=\ceu{\Break}$.
      %%
      Similar to Case~\ref{lem.x.rank-nst.seq-brk}.
      %%
    \item\label{lem.x.rank-nst.loop-adv}$p_1\ne\ceu{\Nop},\ceu{\Break}$.
      %%
      Then~$\pi$ is an instance of~\R{loop-adv}.  Hence there is a
      derivation~$\pi'$ such that
      \[
        \pi'\Vdash\<p_1,n,\nil>\nst\<p_1',n,e_1'>\,,
      \]
      for some~$p_1'$ and~$e_1'$.  Thus~$p'=\ceu{p_1'\AtLoop{\,p_2}}$
      and~$e'=e_1'$.
      %%
      % By the induction hypothesis,
      % \begin{equation}
      %   \label{lem.x.rank-nst.loop-adv-eq1}
      %   \rank(\<p_1,n,\nil>)\ge\rank(\<p_1',n,e_1'>)\,.
      % \end{equation}
      %%
      There are two subcases.
      \begin{case}
      \item$\pot'(p)=\pot'(p_1)$.
        %%
        Then every execution path of~$p_1$ contains a~$\ceu{\Break}$
        or~$\ceu{\AwaitExt}$ statement.  A single~$\nst$ cannot terminate
        the loop, since~$p_1\ne\ceu{\Break}$, nor can it consume
        an~$\ceu{\AwaitExt}$, which means that all execution paths in~$p_1'$
        still contain a~$\ceu{\Break}$ or~$\ceu{\AwaitExt}$.
        Hence~$\pot'(p')=\pot'(p_1')$.  The rest of this proof is similar to
        that of Case~\ref{lem.x.rank-nst.seq-adv}.
        %%
      \item$\pot'(p)=\pot'(p_1)+\pot'(p_2)$.
        %%
        Then some execution path in~$p_1$ does not contain a~$\ceu{\Break}$
        or~$\ceu{\AwaitExt}$ statement.  Since~$p_1\ne\ceu{\Nop}$, a
        single~$\nst$ cannot restart the loop, which means that~$p_1'$ still
        contain some execution path in which a~$\ceu{\Break}$
        or~$\ceu{\AwaitExt}$ does not occur.
        Hence~$\pot'(p')=\pot'(p_1')+\pot'(p_2)$.  The rest of this proof is
        similar to that of Case~\ref{lem.x.rank-nst.seq-adv}.
      \end{case}
    \end{case}

>>>>

\lemxdetout*
\begin{proof}
  The lemma is vacuously true if~$\delta$ cannot be advanced by~$\out$
  transitions.  Suppose that is not the case and let~$\delta=\<p,n,e>$,
  $\delta_1=\<p_1,n_1,e_1>$ and~$\delta_2=\<p_2,n_2,e_2>$.  Then, there are
  two possibilities.
  \begin{case}
  \item $e\ne\nil$.  Both transitions are applications of~\R{push}.
    Hence $p_1=p_2=\bcast(p,e)$, $n_1=n_2=n+1$, and~$e_1=e_2=\nil$.
    %%
  \item $e=\nil$.  Both transitions are applications of~\R{pop}.
    Hence $p_1=p_2=p$, $n_1=n_2=n-1$, and~$e_1=e_2=\nil$.\qedhere
  \end{case}
\end{proof}

@@

    Hence $p_1=p_2=\ceu{\Nop}$ and~$e_1=e_2=\nil$.

@@

    Hence $p_1=p_2=\ceu{\CanRun(n)}$ and~$e_1=e_2=e'$.

@@

      Hence $p_1=p_2=p'$ and~$e_1=e_2=\nil$.

@@

      Hence $p_1=p_2=p''$ and~$e_1=e_2=\nil$.

@@

      Hence $p_1=p_2=\ceu{\Break}$ and~$e_1=e_2=\nil$.

@@

      Hence $p_1=\ceu{p_1';p''}=\ceu{p_2';p''}=p_2$

@@

    Hence $p_1=p_2=\ceu{p'\AtLoop{p'}}$ and~$e_1=e_2=\nil$.

@@

      Hence $p_1=p_2=\ceu{\Loop{p''}}$ and~$e_1=e_2=\nil$.

@@

      Hence $p_1=\ceu{p_1'\AtLoop{p''}}=\ceu{p_2'\AtLoop{p''}}=p_2$

@@

    Hence $p_1=p_2=\ceu{{p'}\AtAnd{(\CanRun(n);\,p'')}}$

@@

        Hence $p_1=p_2=\ceu{\clear(p'');\Break}$ and~$e_1=e_2=\nil$.

@@

        Hence $p_1=\ceu{{p_1'}\And{p''}}=\ceu{{p_2'}\And{p''}}=p_2$

@@

        Hence $p_1=p_2=\ceu{\clear(p');\Break}$ and~$e_1=e_2=\nil$.

@@

        Hence $p_1=\ceu{{p'}\And{p_1''}}=\ceu{{p'}\And{p_2''}}=p_2$

@@

    Hence $p_1=p_2=\ceu{{p'}\AtOr{(\CanRun(n);\,p'')}}$

@@

        Hence $p_1=p_2=\clear(p'')$ and~$e_1=e_2=\nil$.

@@

        Hence $p_1=p_2=\clear(p')$ and~$e_1=e_2=\nil$.

@@

  \item $p=\ceu{p';\,p''}$.
    %%
    There are three subcases.
    \begin{case}
    \item\label{thm.x.term-nst-*.seq-nop} $p'=\ceu{\Nop}$.
      %%
      Then, by~\R{seq-nop} and by the induction hypothesis, there is
      a~$\delta'$ such that
      \[
        \<\ceu{\Nop;\,p''},n,\nil>
        \nst[1]\<p'',n,e>\nst[*]\delta'_\Hnst\,.
      \]
      %%
    \item\label{thm.x.term-nst-*.seq-brk} $p'=\ceu{\Break}$.
      %%
      Then, by~\R{seq-brk},
      \[
        \<\ceu{\Break;\,p''},n,\nil>\nst[1]\<\ceu{\Break},n,\nil>_\Hnst\,.
      \]
      %%
    \item\label{thm.x.term-nst-*.seq-adv}
      $p'\ne\ceu{\Nop},\ceu{\Break}$.
      %%
      Then, by the induction hypothesis, there are~$p_1'$ and~$e$ such that
      \[
        \<p',n,\nil>\nst[*]\<p_1',n,e>_\Hnst\,.
      \]
      By item~\ref{lem.x.props-nst-i.a} of Lemma~\ref{lem.x.props-nst-i},
      \begin{equation}
        \label{thm.x.term-nst-*.seq-adv.eq1}
        \<\ceu{p';\,p''},n,\nil>\nst[*]\<\ceu{p_1';\,p''},n,e>\,.
      \end{equation}
      It remains to be shown that~$\<\ceu{p_1';\,p''},n,e>$ is
      nested-irreducible.  There are four possibilities following from the
      fact that the simpler~$\<p_1',n,e>$ is nested-irreducible.
      %%
      \begin{case}
      \item $e\ne\nil$.  Then, by the definition of~$\Hnst$,
        description~$\<\ceu{p_1';\,p''},n,e>$ is nested-irreducible.
        %%
      \item $p_1'=\ceu{\Nop}$.
        %%
        From~\eqref{thm.x.term-nst-*.seq-adv.eq1},
        \[
          \<\ceu{p';\,p''},n,\nil>\nst[*]\<\ceu{\Nop;\,p''},n,e>\,.
        \]
        From this point on, this case is similar to
        Case~\ref{thm.x.term-nst-*.seq-nop}.
        %%
      \item $p_1'=\ceu{\Break}$.
        %%
        From~\eqref{thm.x.term-nst-*.seq-adv.eq1},
        \[
          \<\ceu{p';\,p''},n,\nil>\nst[*]\<\ceu{\Break;\,p''},n,e>\,.
        \]
        From this point on, this case is similar to
        Case~\ref{thm.x.term-nst-*.seq-brk}.
        %%
      \item$\isblocked(p_1',n)$.
        %%
        Then, by definition,
        \[
          \isblocked(\ceu{p_1';p''},n)=\isblocked(p_1',n)=\mathit{true}\,.
        \]
        Hence from~\eqref{thm.x.term-nst-*.seq-adv.eq1} and by the
        definition~$\Hnst$, description~$\<\ceu{p_1';\,p''},n,e>$ is
        nested-irreducible.
      \end{case}
    \end{case}

  \item\label{thm.x.term-nst-*.loop}
    $p=\ceu{\Loop{p'}}$.
    %%
    Then, by item~\ref{ass.x.syn-rest.loop} of
    Assumption~\ref{ass.x.syn-rest},
    \begin{equation}\label{thm.x.term-nst-*.loop-expd.eq1}
      \<\ceu{\Loop{p'}},n,\nil>\nst[*]\<p_1',n,e>\,,
    \end{equation}
    for some~$e$ and~$p_1'$ such that either~$p_1'=\ceu{\Break\AtLoop{p'}}$
    or~$\isblocked(p_1',n)$.
    \begin{case}
    \item$p_1'=\ceu{\Break\AtLoop{p'}}$.
      %%
      From~\eqref{thm.x.term-nst-*.loop-expd.eq1}, by~\R{loop-brk},
      \begin{align*}
        \<\ceu{\Loop{p'}},n,\nil>
        &\nst[*]\<\ceu{\Break\AtLoop{p'}},n,e>\\
        &\nst[1]\<\ceu{\Nop},n,e>_\Hnst\,.
      \end{align*}
      %%
    \item$\isblocked(p_1',n)$.  Hence
      from~\eqref{thm.x.term-nst-*.loop-expd.eq1} and by the definition
      of~$\Hnst$, $\<p_1',n,e>_\Hnst$.
    \end{case}

@@

\lemxrankout*
\begin{proof}\strut
  Let~$\delta=\<p,n,e>$, $\delta'=\<p',n',e'>$,
  $\rank(\delta)=\<i,j>$, and~$\rank(\delta')=\<i',j'>$.
  \begin{enumerate:a}
  \item Suppose~$\<p,n,e>\outpush\<p',n',e'>$.  Then, by~\R{push},
    $e\ne\nil$, $p'=\bcast(p,e)$, $n'=n+1$, and~$e'=\nil$.  By
    Definition~\ref{def.x.rank}, $j=n+1$, as~$e\ne\nil$, and~$j'=n+1$,
    as~$e'=\nil$ and~$n'=n+1$; hence $j=j'$.
    %%
    It remains to be shown that~$i=i'$:
    \begin{align*}
      i&=\pot(p,e)
         \tag*{by Definition~\ref{def.x.rank}}\\
       &=\pot'(\bcast(p,e))
         \tag*{by Definition~\ref{def.x.pot}}\\
       &=\pot'(p')
         \tag*{since~$p'=\bcast(p,e)$}\\
       &=\pot'(\bcast(p',\nil))
         \tag*{by definition of~$\bcast$}\\
       &=\pot'(\bcast(p',e'))
         \tag*{since~$e'=\nil$}\\
       &=\pot(p',e')
         \tag*{by Definition~\ref{def.x.pot}}\\
       &=i'
         \tag*{by Definition~\ref{def.x.rank}}
    \end{align*}
    Therefore, $\<i,j>=\<i',j'>$.

  \item Suppose~$\<p,n,e>\outpop\<p',n',e'>$.  Then, by~\R{pop}, $p=p'$,
    $n>0$, $n'=n-1$, and~$e=e'=\nil$.
    %%
    By Definition~\ref{def.x.pot}, $\pot(\bcast(p,e))=\pot(\bcast(p',e'))$;
    hence $i=i'$.  And by Definition~\ref{def.x.rank}, $j=n$, as~$e=\nil$,
    and~$j'=n-1$, as~$e'=\nil$ and~$n'=n-1$; hence $j>j'$.
    Therefore, ${\<i,j>}>{\<i',j'>}$.\qedhere
  \end{enumerate:a}
\end{proof}

\lemxranknst*
\begin{proof}
  We proceed by induction on the structure of~$\nst$ derivations.
  Let~$\delta=\<p,n,e>$, $\delta'=\<p',n',e'>$, $\rank(\delta)=\<i,j>$,
  and~$\rank(\delta')=\<i',j'>$.  By the hypothesis of the lemma, there is
  a derivation~$\pi$ such that
  \[
    \pi\Vdash\<p,n,e>\nst\<p',n',e'>\,.
  \]
  By definition of~$\nst$, $e=\nil$ and $n=n'$.  Depending on the structure
  of program~$p$, there are~11 possibilities.  In each one of them we show
  that~$\rank(\delta)\ge\rank(\delta')$.
  \begin{case}
  \item$p=\ceu{\Mem(id)}$.
    %%
    Then~$\pi$ is an instance of~\R{mem}.  Hence $p'=\ceu{\Nop}$
    and~$e'=\nil$.  Thus $\rank(\delta)=\rank(\delta')=\<0,n>$.

  \item$p=\ceu{\EmitInt(e_1)}$.
    %%
    Then~$\pi$ is an instance of~\R{emit-int}.  Hence $p'=\ceu{\CanRun}$
    and~$e'=e_1\ne\nil$.
    Thus
    \[
      {\rank(\delta)={\<1,n>}}>{\<0,n+1>=\rank(\delta')}\,.
    \]

  \item$p=\ceu{\CanRun(n)}$.
    %%
    Then~$\pi$ is an instance of~\R{can-run}.  Hence $p'=\ceu{\Nop}$
    and~$e'=\nil$.  Thus
    \[
      \rank(\delta)=\rank(\delta')=\<0,n>\,.
    \]

  \item$p=\ceu{\IfElse{p}{p_1}{p_2}}$.
    %%
    There are two subcases.
    \begin{case}
    \item\label{lem.x.rank-nst.if-true}$\eval(\ceu{\Mem(\Id)})$.
      %%
      Then~$\pi$ is an instance of~\R{if-true}.  Hence $p'=\ceu{p_1}$
      and~$e'=\nil$.  Thus
      \begin{align*}
        \rank(\delta)&=\<\max\{pot'(p_1),pot'(p_2)\},n>\\
                     &\ge\<\pot'(p_1),n>=\rank(\delta')\,.
      \end{align*}
      %%
    \item$\lnot\eval(\ceu{\Mem(\Id)})$.
      %%
      Similar to Case~\ref{lem.x.rank-nst.if-true}.
    \end{case}

  \item$p=\ceu{p_1;\,p_2}$.
    %%
    There are three subcases.
    \begin{case}
    \item\label{lem.x.rank-nst.seq-nop}$p_1=\ceu{\Nop}$.
      %%
      Then~$\pi$ is an instance of~\R{seq-nop}.
      Hence $p'=p_2$ and~$e'=\nil$.  Thus
      \begin{align*}
        \rank(\delta)&=\<\pot'(p_1)+\pot'(p_2),n>\\
                     &\ge\<pot'(p_2),n>=\rank(\delta')\,.
      \end{align*}
      %%
    \item\label{lem.x.rank-nst.seq-brk}$p_1=\ceu{\Break}$.
      %%
      Then~$\pi$ is an instance of~\R{seq-brk}.
      Hence $p'=p_1$ and~$e'=\nil$.  Thus
      \[
        \rank(\delta)=\rank(\delta')=\<0,n>\,.
      \]
      %%
    \item\label{lem.x.rank-nst.seq-adv}$p_1\ne\ceu{\Nop},\ceu{\Break}$.
      %%
      Then~$\pi$ is an instance of~\R{seq-adv}.  Hence there is a
      derivation~$\pi'$ such that
      \[
        \pi'\Vdash\<p_1,n,\nil>\nst\<p_1',n,e_1'>\,,
      \]
      for some~$p_1'$ and~$e_1'$.  Thus~$p'=p_1';p_2$ and~$e'=e_1'$.  By the
      induction hypothesis,
      \begin{equation}
        \label{lem.x.rank-nst.seq-adv.eq1}
        \rank(\<p_1,n,\nil>)\ge\rank(\<p_1',n,e_1'>)\,.
      \end{equation}
      There are two subcases.
      \begin{case}
      \item$e'=\nil$
        %%
        Then
        \begin{align*}
          \rank(\delta)&=\<\pot'(p_1)+\pot'(p_2),n>\enspace\text{and}\\
          \rank(\delta')&=\<\pot'(p_1')+\pot'(p_2),n>\,.
        \end{align*}
        By~\eqref{lem.x.rank-nst.seq-adv.eq1}, $\pot'(p_1)\ge\pot'(p_1')$.
        Thus
        \[
          \rank(\delta)\ge\rank(\delta')\,.
        \]
        %%
      \item$e'\ne\nil$.
        %%
        Then~$\pi'$ contains one application of~\R{emit-int}, which consumes
        one~$\ceu{\EmitInt(e')}$ statement from~$p_1$ and implies
        $\pot'(p_1)>\pot'(p_1')$.  Thus
        \begin{align*}
          \rank(\delta)&=\<\pot'(p_1)+\pot'(p_2),n>\\
                       &>\<\pot'(p_1')+\pot'(p_2),n+1>=\rank(\delta')\,.
        \end{align*}
      \end{case}
    \end{case}

  \item\label{lem.x.rank-nst.loop-expd}$p=\ceu{\Loop{p_1}}$.
    %%
    Then~$\pi$ is an instance of~\R{loop-expd}.
    Hence $p'=\ceu{p_1\AtLoop{p_1}}$ and~$e'=\nil$.
    %%
    By item~\ref{ass.x.syn-rest.loop} of Assumption~\ref{ass.x.syn-rest},
    all execution paths of~$p_1$ contain at least one occurrence
    of~$\ceu{\Break}$ or~$\ceu{\AwaitExt}$.  Thus, by condition~(\dag) in
    Definition~\ref{def.x.pot},
    \[
      \rank(\delta)=\rank(\delta')=\<\pot'(p_1),n>\,.
    \]

  \item$p=\ceu{{p_1}\AtLoop{p_2}}$.
    %%
    There are three subcases.
    \begin{case}
    \item$p_1=\ceu{\Nop}$.
      %%
      Similar to Case~\ref{lem.x.rank-nst.seq-nop}.
      %%
    \item$p_1=\ceu{\Break}$.
      %%
      Similar to Case~\ref{lem.x.rank-nst.seq-brk}.
      %%
    \item\label{lem.x.rank-nst.loop-adv}$p_1\ne\ceu{\Nop},\ceu{\Break}$.
      %%
      Then~$\pi$ is an instance of~\R{loop-adv}.  Hence there is a
      derivation~$\pi'$ such that
      \[
        \pi'\Vdash\<p_1,n,\nil>\nst\<p_1',n,e_1'>\,,
      \]
      for some~$p_1'$ and~$e_1'$.  Thus~$p'=\ceu{p_1'\AtLoop{\,p_2}}$
      and~$e'=e_1'$.
      %%
      % By the induction hypothesis,
      % \begin{equation}
      %   \label{lem.x.rank-nst.loop-adv-eq1}
      %   \rank(\<p_1,n,\nil>)\ge\rank(\<p_1',n,e_1'>)\,.
      % \end{equation}
      %%
      There are two subcases.
      \begin{case}
      \item$\pot'(p)=\pot'(p_1)$.
        %%
        Then every execution path of~$p_1$ contains a~$\ceu{\Break}$
        or~$\ceu{\AwaitExt}$ statement.  A single~$\nst$ cannot terminate
        the loop, since~$p_1\ne\ceu{\Break}$, nor can it consume
        an~$\ceu{\AwaitExt}$, which means that all execution paths in~$p_1'$
        still contain a~$\ceu{\Break}$ or~$\ceu{\AwaitExt}$.
        Hence $\pot'(p')=\pot'(p_1')$.  The rest of this proof is similar to
        that of Case~\ref{lem.x.rank-nst.seq-adv}.
        %%
      \item$\pot'(p)=\pot'(p_1)+\pot'(p_2)$.
        %%
        Then some execution path in~$p_1$ does not contain a~$\ceu{\Break}$
        or~$\ceu{\AwaitExt}$ statement.  Since~$p_1\ne\ceu{\Nop}$, a
        single~$\nst$ cannot restart the loop, which means that~$p_1'$ still
        contain some execution path in which a~$\ceu{\Break}$
        or~$\ceu{\AwaitExt}$ does not occur.
        Hence $\pot'(p')=\pot'(p_1')+\pot'(p_2)$.  The rest of this proof is
        similar to that of Case~\ref{lem.x.rank-nst.seq-adv}.
      \end{case}
    \end{case}

  \item\label{lem.x.rank-nst.and-expd}$p=\ceu{{p_1}\And{p_2}}$.
    %%
    Then~$\pi$ is an instance of~\R{and-expd}.
    Hence $p'=\ceu{{p_1}\AtAnd(\CanRun(n);\,p_2)}$ and~$e'=\nil$.
    Thus
    \[
      \rank(\delta)=\rank(\delta')=\<\pot'(p_1)+\pot'(p_2),n>\,.
    \]
